---
title: "NBA PLAYERS SALARY ANALYSIS"
author: "Kevin Andrian Suryanto"
output: html_notebook
---

#Abstract
A common person would assume that the success of an NBA player's career is directly correlated with how much points they can score, and therefore would earn more salary if they have more points. However, points scored by an NBA player typically aren't the only measure of success. Moreover, although some NBA players are considered more successful than others, some of those players identified as less successful earn more than some players that are indentified as more successful.

#Research Question
What affects an NBA player's salary? What are the biggest contributing factors to a player's salary? Does more playing time result in higher salary? Does different players positions affect salary?

#Packages needed
```{r}
library(magrittr)
library(dplyr)
library(tidyverse)
library(DataComputing)
```

#Data for Analysis
```{r}
nbastats <- read.csv("nba_2017_players_stats_combined.csv")
nbasalary <- read.csv("nba_2017_salary.csv")
```
These 2 datasets are obtained from Kaggle (https://www.kaggle.com/noahgift/social-power-nba), an online community of data scientists and machine learning practitioners. The data is collected by the NBA and uploaded by Noah Gift in 2017. This data is used in several studies for the IBM Developerworks: "Explore valuation and attendance using data science and machine learning" and "Exploring the individual NBA players."

###Variables
* Rk = Player's rank in terms of points per game
* PLAYER = Player's name
* POSITION = Player's position
* AGE = Player's age
* MP = Minutes played
* FG = Field goals made per game
* FGA = Field goals attempted per game
* FG. = Field goal % (made over attempted)
* X3P = 3-point field goals made per game
* X3PA = 3-point field goals attempted per game
* X3P. = 3-point field goal % (made over attempted)
* X2P = 2-point field goals made per game
* X2PA = 2-point field goals attempted per game
* X2P. = 2-point field goal % (made over attempted)
* eFG. = Effective field goal %
* FT = Free throws made per game
* FTA = Free throws attempted per game
* FT. = Free throw % (made over attempted)
* ORB = Offensive rebounds per game
* DRB = Defensive rebounds per game
* TRB = Total rebounds per game
* STL = Steals per game
* BLK = Blocks per game
* TOV = Turnovers per game
* PF = Personal fouls committed per game
* Points = Points per game
* TEAM = Player's team abbreviation
* GP = Number of games played
* MPG = Minutes played per game]
* ORPM = Offensive real plus-minus per game
* DRPM = Defensive real plus-minus per game
* RPM = Real plus-minus per game
* WINS_RPM = An estimate of the number of wins a player has contributed to his team's win total on the season
* PIE = Player's impact estimate (a player's impact on each individual game they play)
* PACE = Estimate of the number of possessions per 48 minutes
* W = Number of wins
* SALARY = Amount of yearly salary (in $)

###Glimpse of Data
```{r}
head(nbastats)
head(nbasalary)
```
The variable "POSITION" and "TEAM" will be deleted from nbasalary since "POSITION" and "TEAM" are already present in the nbastats dataset. The first colum will be deleted from nbastats since it is unnecessary. The two datasets will be joined by the player's name. Duplicates of NBA players (if any) will also be removed.
```{r message=FALSE, warning=FALSE}
nbastats <- nbastats[,-1]
nbasalary <- nbasalary[,-2][,-2]
nba <- nbastats %>%
  left_join(nbasalary, by = c("PLAYER" = "NAME"))
nba = nba[!duplicated(nba$PLAYER),]
head(nba)
```

###Data Description
The dataset "nbastats"" shows statistics for NBA players' on-court performances in the 2016-2017 season. The dataset "nbasalary" shows NBA players' salaries in the 2016-2017 season. The combined dataset has 446 cases. Each case represents an NBA player.
```{r}
nrow(nba)
```
#How much does an NBA player's salaries vary?
```{r}
summary(nba$SALARY)
```
In 2017, the NBA player's salary has a minimum of $31,696 (Pierre Jackson), a maximum of $30,963,450 (LeBron James), and a mean of $7,415,807.
Here are the 5 highest and lowest earning NBA players.
```{r}
attach(nba)
#Top 5 highest earning players
nba %>%
  select(PLAYER,SALARY) %>%
  arrange(desc(SALARY)) %>%
  slice(1:5)
#Top 5 lowest earning players
nba %>%
  select(PLAYER,SALARY) %>%
  arrange(SALARY) %>%
  slice(1:5)
```

#Does an NBA player's points result in higher salary?
```{r}
ggplot(data=nba,aes(x=POINTS,y=SALARY))+geom_point() + stat_smooth(method=lm) + ggtitle("Salary of NBA players in terms of PPG (figure 1.1)")
```
When referring to figure 1.1, there seems to be an upward trend which shows that higher points per game typically results in higher salary although it's not for all cases.
```{r}
nba %>% 
  filter(grepl("[F|C]", POSITION, ignore.case = TRUE)) %>%
  group_by(POSITION) %>%
  summarise(totalpoints=sum(POINTS))
```
In the bigman positions(forwards and centers), each position seem to have a similar number of total points except PF-C.
```{r}
nba %>% 
  filter(grepl("[G]", POSITION, ignore.case = TRUE)) %>%
  group_by(POSITION) %>%
  summarise(totalpoints=sum(POINTS))
```
In the smaller positions(guards), both positions only have a difference of 80.6 total points and are also higher than bigman positions on average.
```{r}
ggplot(data=nba,aes(x=POSITION,y=SALARY ))+geom_bar(stat='identity',position='stack', width=.9)+ ggtitle("Salary of NBA players in terms of PPG per position (figure 1.2)")  
```
According to figure 1.2, all 5 positions have similar salaries. However, Centers have the highest salary.
```{r}
ggplot(data=nba,aes(x=POINTS,y=SALARY))+geom_point()+facet_wrap(~POSITION,ncol=4) + stat_smooth(method=lm) + ggtitle("Salary of NBA players in terms of PPG per position (figure 1.3)") 
```
According to figure 1.3, higher points result in higher salary amongst all player positions. Furthermore, we will observe how other key NBA stats affect a player's salary.

## Assists per game
```{r}
ggplot(data=nba,aes(x=AST,y=SALARY))+geom_point() + stat_smooth(method=lm) + ggtitle("Salary of NBA players in terms of APG (figure 2.1)")
```
We can obseve that there is an upward trend in salary in terms of assists per game.

## Rebounds per game
```{r}
ggplot(data=nba,aes(x=TRB,y=SALARY))+geom_point() + stat_smooth(method=lm) + ggtitle("Salary of NBA players in terms of RPG (figure 2.2)")
```
We can obseve that there is an upward trend in salary in terms of rebounds per game.

## Steals per game
```{r}
ggplot(data=nba,aes(x=STL,y=SALARY))+geom_point() + stat_smooth(method=lm) + ggtitle("Salary of NBA players in terms of SPG (figure 2.3)")
```
We can obseve that there is a slight upward trend in salary in terms of steals per game although it is not significant.

## Blocks per game
```{r}
ggplot(data=nba,aes(x=BLK,y=SALARY))+geom_point() + stat_smooth(method=lm) + ggtitle("Salary of NBA players in terms of BPG (figure 2.4)")
```
We can see that there is not much trend between blocks per game and a player's salary. Out of 5 of the players with 2 or more blocks per game, 4 made a salary of lower than average.

## Minutes played per game
```{r}
ggplot(data=nba,aes(x=MP,y=SALARY))+geom_point() + stat_smooth(method=lm) + ggtitle("Salary of NBA players in terms of MPG (figure 2.5)")
```
We can obseve that there is an upward trend in salary in terms of minutes played per game.

However, although all of these key NBA stats graphically show an increasing trend, we will further perform a simple linear regression test the significance of these NBA statistic measures.

##Simple Linear Regression
Linear regression models are used to show or predict the relationship between two variables or factors. We will fit salary as the response variable and use number of points, assists, rebounds, steals, blocks, minutes, and position as the predictors.
```{r}
model <- lm(SALARY~POINTS+AST+TRB+STL+BLK+MP)
summary(model)
```
From the p-value results above, we are 95% confident that number of points, assist, and rebound have a significant relationship with a player's salary. However, steals, blocks, and minutes player does not significantly affect a player's salary since their p-values are larger than 0.05.

#Conclusion
From the graphical and linear regression observations, we can conclude that the number of points per game by an NBA player significantly affects an NBA player's salary. Moreover, other key important factors such as number of assists and rebounds also significantly affect a player's salary. Some people may say that more playing time (or minutes played) have higher salaries. However, although our graph shows that there is a positive relationship between minutes played and salary, the linear regression output shows that the effect of minutes played is not significant to salary. From the salary graphs in terms of positions, different positions does not show significance in salary difference. 
